Seasonality of submesoscale dynamics in the Kuroshio Extension

### 2016-09-22 writeup/RochatEtal_ResponseToReviewers.tex

%\include{symbols}

\maketitle

\section*{Summary}

In this revised manuscript, we have addressed the reviewers'
comments. Fundamentally, we have included a new wavenumber-frequency
spectral analysis suggested by reviewer 3. The new results vindicate
our claim that inertia-gravity waves, particularly super-inertial waves,
account for most of the summertime SSH variance and surface KE at scales
between 10 and 100 km. We have also expanded the introduction and discussion
to place our study in the context of Surface Water \& Ocean Topography (SWOT)
satellite mission.

@@

\subsection{Major points}

\begin{enumerate}

  \item {\it First of all, generation mechanism of submesoscale inertial-gravity waves in summer
        and early fall is unclear to me. In the paper, authors attribute the mechanism to
        internal tides (X-12, section 6, last sentence) and emphasize using the tidal forcing in
        their model (X-5, section 2, 2nd paragraph). However, authors do not explain or show
        any interaction mechanism of internal tides (please at least add some reference). With
        the fixed frequencies of tidal constituents and existence of the critical latitude for the
        diurnal tides, I am not sure what mechanism can represent the surface near-inertial
        wave generation within the whole analyzed region by internal tides (with the
        horizontal scales of 10-100km). If authors compare Figures 2 and 4 with those
        estimated from the non-tidal forcing run in the Supplemental material, authors'
        conclusion could be strengthened.}\\

        \bdp
          The generation mechanism for internal
          waves involves changes in the wind (NIWs) and barotropic tides sloshing over
          topographic features --- we are not claiming that there exists a different
          generation mechanism
          in summer compared to winter; neither are we claiming that surface NIWs
          are generated by internal tides. What is different is the near-surface
          expression of IGWs, which is very sensitive to the near-surface stratification
          \citep[; cf. second paragraph of discussion]{dasaro1978}.\\

@@

        \bdp
            We are aware of the this issue --- the effective resolution based on the KE spectrum
            was already mentioned
            in the supporting information to the first draft.  In
            fact, the smallest scale discussed in our paper (10 km) was set by the effective
            resolution of highest-resolution simulation. To be more explicit, we have now included
            the effective resolution together with the nominal resolutions in the main text (Section 2).\\

            The ``unresolved energy'' does not come from anywhere --- flows smaller than
            about 10 km are strongly damped.
            To the extent that 1-km-scale flows are associated with a forward
            energy transfer, then their effects on the scales concerning our study
            (10-100 km) are likely energetically insignificant.
            Also important to this discussion is that higher resolution simulations will resolve {\it both}
            1-km-scale eddies and 1-km-scale IGWs that are generated via wave-wave
            interactions and wave-vorticity interactions.\\

            \cite{buckingham_etal2016} made a good point that models and observations
            to date may not resolve the mixed-layer deformation scale in summer; this is no
            different for our simulations. But the interpretation that this is the main
            source of seasonally varying eddy kinetic energy is unclear because
            summer is also the season when lateral buoyancy
            gradients (and thus available potential energy) are minimum within the mixed layer
            \citep[e.g., ][]{callies_etal2015}.\\

@@

  \item {\it (Section 1, last paragraph) Adding “with implications for SWOT” at the end of the
          introduction would also help this article.}

          \bdp
            Added.
          \edp

  \item {\it (Section 1, Paragraph 1) The authors mention a suite of papers documenting
        seasonality at the submesoscale but fail missed a few: Ostrovskii (1995) and
        Brannigan et al. (2015). These are observation- and model-based studies, resp.}\\

        We now reference to the recent paper byt Brannigan et al. (2015).
        Ostrovskii (1995) can accessed via Buckingham et al. (2016).

### 2016-09-22 writeup/rocha_etal.tex
Minor changes

\begin{abstract}
    Submesoscale turbulence, fueled by ample potential
    energy stored in wintertime deep mixed layers, undergoes vigorous seasonality.
    Here we provide model evidence that the seasonally-changing upper-ocean
    stratification also modulates submesoscale --- here 10-100 km ---
    inertia-gravity waves. Summertime re-stratification weakens submesoscale
    turbulence but enhances inertia-gravity waves near the surface. Thus,
    submesoscale turbulence and inertia-gravity waves undergo vigorous out-of-phase
    seasonal cycles. The present results imply a strong seasonal modulation of
    the skill and accuracy of geostrophic velocity diagnostics from submesoscale
    sea-surface height delivered by the Surface and Water Ocean Topography (SWOT)
    satellite mission
\end{abstract}

@@

% now SWOT
The partition between geostrophic and ageostrophic flows across
the submesoscales and their seasonality have immediate applications for the planning
of the Surface and Water Ocean Topography (SWOT) satellite mission.
SWOT will provide
sea-surface height (SSH) with about 15 km resolution, thereby providing the
first global SSH measurements at submesoscales \citep{fu_ubelmann2014}. A
challenge --- and opportunity ---  for SWOT is the presence of ageostrophic flows
that straddle the submesoscales and project
onto SSH \citep[e.g., ][]{richman_etal2012} thus making submesoscale geostrophic
velocities difficult to extract from SSH.

Using the output of $1/24^\circ$ and $1/48^\circ$ global
numerical simulations with embedded tides, we show that IGWs undergo
a strong seasonality near the surface in the Kuroshio Extension region.
Interestingly, the seasonal cycle of IGWs is out of phase
with the seasonal cycle of submesoscale turbulence. Consistent with previous studies,
deep late-winter mixed layers are prone to
shallow baroclinic instabilities that are roughly in geostrophic balance
and flux energy upscale \citep{sasaki_etal2014,callies_etal2016},
driving a mild seasonal modulation of the mesoscales \citep{sasaki_etal2014,qiu_etal2014}.

IGWs with horizontal scales between 10-100 km (hereafter submesoscale IGWs),
however, peak in late summer/early fall,
when the upper ocean is strongly stratified. Thus there exists a
strong seasonal modulation of upper-ocean submesoscale dynamics:
submesoscale turbulence dominates the upper-ocean dynamics in late winter/early fall,
whereas submesoscale IGWs prevail in late summer/early fall.
Because submesoscale turbulence is weakest in late summer/early fall,
IGWs account for most of the
summertime submesoscale sea-surface height variability with implications for SWOT.

\section{The LLC numerical simulations}
We use results of two latitude-longitude polar cap (LLC)
realistic numerical simulations. The outputs
analyzed here, LLC2160 (nominal resolution 1/24$^\circ$$\approx$4.7 km;
effective resolution $\approx$ 20 km) and LLC4320
(nominal resolution 1/48$^\circ$$\approx$2.3 km;
effective resolution $\approx$8 km),   are
forward solution of the Massachussetts Institute of Technology general circulation model \citep[MITgcm; ][]{marshall_etal1997}
on a LLC grid  \citep{forget_etal2015} with
90 vertical levels. The coarser-resolution LLC simulation was
spun up from an Estimating the Circulation and Climate of the Ocean, Phase II \citep[ECCO2; ][]{menemenlis_etal2008}
adjoint-method state estimate, constrained to millions
of observations from 2009 through 2011. Both simulations were forced by
tides hourly and by 6-hourly surface atmospheric fields. The LLC2160
output spans two years from March 2011 to April 2013; the LLC4320 was spun up from
the LLC2160 simulation, spanning one year from September 2011 to October 2012.
The LLC4320 simulation is an extension of the 3-month long output used by
\citet{rocha_etal2016} --- for details, see supporting information.

A key aspect of the LLC2160 and LLC4320 simulations is that they were forced by
the 16 most-significant tidal components.
Because barotropic tides interact with topography and generate internal
tides that project onto mesoscales to submesoscales
\citep[e.g., ][]{rocha_etal2016}, tidal forcing fundamentally distinguishes our analysis
from other modeling studies of seasonality \citep{sasaki_etal2014,qiu_etal2014}.

To study seasonal variations in the upper-ocean dynamics, we focus on the northwest
Pacific, in the vicinity of the Kuroshio
Extension, where previous studies have shown strong mesoscale and submesoscale seasonality
\citep{sasaki_etal2014,qiu_etal2014}.
We analyze a sub-domain of the LLC4320 and LLC2160 simulations of about 2000 km$^2$
spanning 155-175$^\circ$E; 25-40$^\circ$N (Figure \ref{fig1}a). The stratification
in this mesoscale-rich subtropical region undergoes a vigorous seasonal cycle: wintertime-enhanced
small-scale turbulence de-stratifies the upper ocean, yielding mixed layers
deeper than 200 m to the north (Figure \ref{fig1}b). Summertime re-stratification yields mixed layers shallower
than 40 m (Figure \ref{fig1}c).
Fundamentally, the upper-ocean density structure is well-captured by both LLC simulations:
a comparison with Argo climatology shows that both simulations skillfully represent the Kuroshio
Extension stratification and its seasonal variability  (supporting information). Moreover, model smoothed
fields have eddy kinetic energy (EKE) levels ($0.041$ m$^2$ s$^{-2}$) similar to Jason II across-track
geostrophic EKE ($0.036$ m$^2$ s$^{-2}$) and gridded geostrophic EKE ($0.032$ m$^2$ s$^{-2}$).

@@

\section{Bulk statistics of the surface lateral velocity gradient tensor}
To study the seasonality in the surface velocity, we calculate the lateral velocity gradient tensor
\begin{equation}
\left[\begin{matrix} u_x & u_y\\ v_x&v_y \end{matrix}\right]
\end{equation}
using a centered
second-order finite difference scheme --- we obtain qualitatively similar results
using spectral differentiation of velocity fields periodized with reflections;
statistics obtained via spectral differentiation are 10-15$\%$ larger.
We then diagnose
the vertical vorticity
\begin{equation}
\zeta \equiv v_x - u_y\, ,
\end{equation}
lateral rate of strain
\begin{equation}
  \alpha \equiv [(u_x-v_y)^2 + (u_y + v_x)^2]^{1/2}\,,\\
\end{equation}
and horizontal divergence
\begin{equation}
\delta \equiv u_x + v_y\, .
\end{equation}
Because vortices, fronts, and IGWs are ubiquitous at submesoscales, second-order
statistics of vorticity, strain, and divergence are apt descriptors of these
fine lateral scales \citep[e.g.,][]{shcherbina_etal2013}.

Figures \ref{fig1}b-c show snapshots of vertical vorticity $\zeta$ in early spring
(April 15) and fall (October 15) in the LLC4320 (1/48$^\circ$) simulation.
The model solutions depict seasonality in vorticity: large values of
fine-grained vertical vorticity are observed in early spring with maximum
values as large as $4f$, where $f$ is
the local planetary vorticity, and root-mean-square (RMS) of about $0.4f$. In early
fall, the situation is the opposite: the vertical vorticity is relatively coarse-grained;
its local maximum and RMS are both smaller than $0.5f$.
Indeed, the vorticity and rate of strain are strongest in wintertime (Figure \ref{fig2}):
in both simulations, the RMS vorticity and strain rate are about twice as large in
late winter/early spring
than in late summer/early fall. Because the wintertime vorticity and strain rate
are dominated by the smallest scales in the flow (the KE spectrum is shallower than
a $-3$ power law in winter), increasing the resolution from
1/24$^\circ$ to 1/48$^\circ$ increases the wintertime RMS vorticity
and strain by about 40$\%$.

@@

The bulk of vertical vorticity and strain rate are associated with subinertial flows
($T_{32.5} = 2\pi/f_{32.5}\approx 22.3$h, where $f_{32.5}$ is the inertial frequency at
the  mid-latitude $32.5^\circ$N): daily-averaging the velocity fields suppresses super-inertial
motions and reduces the RMS
vorticity by 40$\%$ and the RMS strain by 10$\%$; the seasonal
cycle remains strong (see
dashed lines in figures \ref{fig2}a-b). Indeed, most of this seasonal cycle is associated
with submesoscale flows: smoothing the velocity fields
with a Hanning filter with cut off scale of 100 km dramatically reduces the RMS
vorticity and strain rate. The reduction in variance is about 80$\%$ in winter,
yielding RMS vorticity and strain  rate roughly consistent with the diagnostics from
AVISO gridded geostrophic velocities (compare red lines to black lines in figures \ref{fig2}a-b).
The picture that emerges is consistent with recent studies: shallow baroclinic
instabilities energize the submesoscales in winter, drawing from the available
potential energy stored in large lateral buoyancy gradients in deep mixed layers \citep{sasaki_etal2014,callies_etal2015,callies_etal2016}.

The seasonal cycle of the horizontal divergence, however, showcases the complexity
of the upper-ocean annual variability. If submesoscale eddies and fronts dominated
the near-surface variability all year, then the seasonal cycle of horizontal divergence,
vertical vorticity, and lateral strain rate would be in phase \citep[e.g.,][]{sasaki_etal2014}.
While there is a clear wintertime peak in divergence of daily-averaged velocity
(see dashed lines in figure \ref{fig2}c; RMS divergence $\sim0.1 f$ in the 1/48$^\circ$
simulation), the hourly fields show
a stronger enhancement of
lateral divergence in late summer/early fall (RMS divergence $\sim0.22 f$ in the 1/48$^\circ$
simulation). Because the 1/48$^\circ$ simulation better resolves
smaller-scale submesoscale flows, a secondary RMS divergence peak in winter  is
nearly as strong as in summer. Submesoscale fronts and eddies evolve
relatively fast, and there is no clear
temporal and spatial scale separation between those motions and IGWs
\citep[][]{mcwilliams2016}:
daily-averaging the velocity fields efficiently suppresses the summertime horizontally
divergent flows, but also reduces the wintertime lateral divergence by about 50$\%$.
Figure \ref{fig2}c also shows that most of the lateral divergence is associated
with submesoscale flows: smoothing the velocity fields with a 100-km-cutoff
suppresses more than 80$\%$ of the RMS divergence.

\begin{figure*}[ht]
  \begin{center}
    \includegraphics[width=.75\textwidth]{figs/fig3.pdf}
 \caption{\small Seasonal variation of joint probability distributions of
          surface lateral velocity gradient tensor:  vorticity vs.\
          strain rate (a through d),
          and vorticity vs.\ Laplacian of sea-surface height (e through h) in April (a, b,
          e, f) and October (c, d, g, h).
          Dashed lines in (a) through (d) represent one dimensional shear flow $\alpha = \pm\zeta$,
          characteristic of fronts. Dashed lines in (e) through
          (h) represent geostrophic flow $\zeta = \frac{g}{f}\nabla_h^2 \eta$.}
 \label{fig3}
 \end{center}
\end{figure*}

@@

The April vorticity-strain joint-PDF has a shape characteristic of submesoscale turbulence
 (Figures \ref{fig3}a-b). The alignment of vorticity and strain $\alpha \sim \pm\zeta$
 with strong positive skewness are fingerprints of submesoscale fronts
 \citep{shcherbina_etal2013,mcwilliams2016}. The shape of the vorticity-strain
 joint-PDF is similar for hourly and daily-averaged fields, although
 the vorticity skewness reduces from 1.4 to 1.13 from hourly to daily-averaged.
 The April results indicate that wintertime submesoscale
 surface velocity is strongly dominated by submesoscale turbulence.
 The hourly velocity and sea-surface height fields have an important
ageostrophic component as depicted by the joint-PDF of vorticity-Laplacian of SSH
winter (Figure \ref{fig3}e). The daily-averaged fields
are largely
in  geostrophic balance (Figure \ref{fig3}f).

@@

Figures \ref{fig4}a-b show the wavenumber-frequency of SSH variance for the LLC4320
simulation. Mesoscales flows and the lunar semi-diurnal tide contain most of the
SSH variance. But there is significant SSH variance at submesoscales, both sub-inertial
and super-inertial. Sub-inertial SSH variance at submesoscales is larger in April, whereas
super-inertial SSH variance at submesoscales is larger in October. At scales smaller
than 100 km (submesoscales), the SSH variance is 66$\%$ larger in October than in April.
IGWs, particularly modes two through four, account for most of the SSH variance
increase in summer --- higher vertical
modes are more sensitive to changes in near-surface stratification (supporting information).
In October, super-inertial IGWs account for a SSH standard deviation of about 4.9 cm,
comparable to the April sub-inertial SSH standard deviation of $\sim$5.2 cm (Table \ref{tab1}).
The
Similarly, the submesoscale super-inertial surface KE is 63$\%$ larger in October
than in April (supporting information).

@@

Figure \ref{fig4}c shows the integral over frequency
of wavenumber-frequency spectra in figures \ref{fig4}a-b.
Both in April and October, the
SSH wavenumber spectrum approximately follows a -2 power-law at submesoscales
(solid lines in figure \ref{fig4}c)  --- there is
small difference between the spectra owing to the phase cancellation between the
seasonal cycle of submesoscale turbulence and inertia-gravity waves. The sub-inertial flow
has steeper submesoscale SSH variance wavenumber spectrum (dashed lines in figure \ref{fig4}c).
 The difference between the SSH variance wavenumber spectrum for all frequencies and
 sub-inertial frequencies is dramatic in October when super-inertial flows account for
 about 80\% of the SSH variance and KE at submesoscales. We obtain similar results by computing
SSH variance and KE wavenumber spectra directly from hourly and daily-averaged
SSH velocity fields.

\begin{figure*}[ht]
  \begin{center}
    \includegraphics[width=.8\textwidth]{figs/fig4.pdf}
 \caption{\small LLC4320 wavenumber-frequency spectrum of SSH variance in (a) April and (b) October.
          (c) Wavenumber spectrum of SSH variance --- the integral of (a) and (b) over frequency.
          In (a) and (b), the light gray shaded region depicts the dispersion relations for inertia-gravity
          waves from mode 1 through mode 4 across the latitudinal domain; horizontal dashed lines represent the semi-diurnal
          lunar tidal frequency ($M_2$), and the inertial frequency at mid-latitude
          ($f_{32.5}$) and its first harmonic (2$f_{32.5}$). }
  \label{fig4}
 \end{center}
\end{figure*}

@@

Internal tide energy flux does not show a seasonal modulation \citep[e.g.,][]{alford2003};
hence one expects a strong seasonality in the near-surface
expression of internal tides and
higher frequency IGWs if the near-surface stratification varies strongly.
Windy near-inertial wave (NIW) generation is intermittent and peaks in early winter, but
NIW energy has a similar energy level in October and April
\citep{alford_etal2016}. Furthermore, in our simulations the higher frequency inertia-gravity-wave
account for most of the seasonality in super-inertial surface KE and SSH
(Figure \ref{fig4}a-b).

The seasonally varying MLD challenges observation and simulation of
flows at ML-deformation-radius length scale in observational programs and models with
fixed resolution \citep{buckingham_etal2016}. But the potential energy stored in the
ML is lowest in summer owing to shallow MLD and weaker lateral buoyancy
gradients \citep{callies_etal2015}. Moreover, in baroclinic turbulence, the deformation
radius has a secondary, catalytic role --- most of the baroclinic convertion occurs at
larger scales \citep{larichev_held1995,callies_etal2016}.
Thus, the model failure to resolve the summertime
ML deformation radius is unlikely to account for the vigorous seasonality in submesoscale
turbulence.

In the context of SWOT, our results suggest that if one is interested
only in the geostrophic flow, so that surface velocity can be diagnosed
from SSH, then the noise-to-signal ratio will have a strong seasonality owing
to the presence of IGWs --- the coherent fraction of internal tides may be removed
efficiently.
An important caveat is that the Kuroshio Extension may be typical of a mesoscale-rich
subtropical regions, but it is unlikely to be
representative of other regions such as low-eddy-kinetic-energy eastern boundary
currents.  We plan to report on the geographic variability of submesoscale
seasonality in a future study.
hing.

>>>>

\begin{abstract}
    The vigorous seasonal cycle of mixed layer modulates upper-ocean submesoscale turbulence.
    Here we provide model evidence that the seasonally-changing upper-ocean
    stratification also modulates submesoscale  (here 10-100 km)
    inertia-gravity waves. Summertime re-stratification weakens submesoscale
    turbulence but enhances inertia-gravity waves near the surface. Thus,
    submesoscale turbulence and inertia-gravity waves undergo vigorous out-of-phase
    seasonal cycles. These results imply a strong seasonal modulation of
    the skill and accuracy of geostrophic velocity diagnosed from submesoscale
    sea-surface height delivered by the Surface and Water Ocean Topography (SWOT)
    satellite mission
\end{abstract}

@@

% now SWOT
The partition between geostrophic and ageostrophic flows across
the submesoscales and their seasonality have immediate applications for the planning
of the Surface and Water Ocean Topography (SWOT) satellite mission.
SWOT will deliver
sea-surface height (SSH) with about 15 km resolution, thereby providing the
first global SSH measurements at submesoscales \citep{fu_ubelmann2014}.
The projection of energetic submesoscale ageostrophic flows, such as
inertia-gravity waves (IGWs), onto SSH represents
a challenged --- and an opportunity --- for SWOT. The challenge: how to diagnose
surface velocity from SSH. The opportunity: observations of submesoscale ageostrophic
flows on global scale.

Using the output of two comprehensive global numerical simulations, here we show
that IGWs undergo
a strong near-surface seasonal cycle in the Kuroshio Extension region.
Interestingly, the seasonal cycle of IGWs is out of phase
with the seasonal cycle of submesoscale turbulence. Consistent with previous studies,
submesoscale turbulence is strongest  in winter \citep{sasaki_etal2014,qiu_etal2014}.

IGWs with horizontal scales between 10-100 km (hereafter submesoscale IGWs),
however, peak in late summer/early fall,
when the upper ocean is strongly stratified. Thus there exists a
strong seasonal modulation of upper-ocean submesoscale dynamics:
submesoscale turbulence dominates the upper-ocean dynamics in late winter/early fall,
whereas submesoscale IGWs prevail in late summer/early fall.

\section{The LLC numerical simulations}
We use results of two latitude-longitude polar cap (LLC)
comprehensive numerical simulations. The outputs
analyzed here, LLC2160 (nominal resolution 1/24$^\circ$$\approx$4.7 km;
effective resolution $\approx$ 20 km) and LLC4320
(nominal resolution 1/48$^\circ$$\approx$2.3 km;
effective resolution $\approx$8 km),   are
forward solution of the Massachussetts Institute of Technology general circulation model \citep[MITgcm; ][]{marshall_etal1997}
on a LLC grid  \citep{forget_etal2015} with
90 vertical levels. The coarser-resolution LLC simulation was
spun up from an Estimating the Circulation and Climate of the Ocean, Phase II \citep[ECCO2; ][]{menemenlis_etal2008}
adjoint-method state estimate, constrained to millions
of observations from 2009 through 2011.

The LLC2160
output spans two years from March 2011 to April 2013; the LLC4320 was spun up from
the LLC2160 simulation, spanning one year from September 2011 to October 2012.
Both LLC2160 and LLC4320 simulations were forced by
tides hourly and by 6-hourly surface atmospheric fields.
The LLC4320 simulation is an extension of the 3-month long output used by
\citet{rocha_etal2016} --- for details, see supporting information.

Tidal forcing by the 16 most-significant components represents a key feature LLC2160 and
LLC4320 simulations. Barotropic tides flowing over topography convert significant tidal energy into the internal
wave field. Those internal waves project onto mesoscales to submesoscales
\citep[e.g., ][]{rocha_etal2016}. Hence tidal forcing fundamentally distinguishes our analysis
from other modeling studies of seasonality \citep{sasaki_etal2014,qiu_etal2014}.

@@

To study seasonal variations in the upper-ocean dynamics, we focus on the northwest
Pacific, in the vicinity of the Kuroshio
Extension, where previous studies have shown strong mesoscale and submesoscale seasonality
\citep{sasaki_etal2014,qiu_etal2014}.
We analyze a sub-domain of the LLC4320 and LLC2160 simulations of about 2000 km$^2$
spanning 155-175$^\circ$E; 25-40$^\circ$N (Figure \ref{fig1}a). The stratification
in this mesoscale-rich subtropical region undergoes a vigorous seasonal cycle: wintertime-enhanced
small-scale turbulence de-stratifies the upper ocean, yielding mixed layers
deeper than 200 m to the north (Figure \ref{fig1}b). Summertime re-stratification yields mixed layers shallower
than 40 m (Figure \ref{fig1}c).

Fundamentally, the upper-ocean density structure is well-captured by both LLC simulations:
a comparison with Argo climatology shows that both simulations skillfully represent the Kuroshio
Extension stratification and its seasonal variability  (supporting information). Moreover, model smoothed
fields have eddy kinetic energy (EKE) levels ($0.041$ m$^2$ s$^{-2}$) similar to Jason II across-track
geostrophic EKE ($0.036$ m$^2$ s$^{-2}$) and gridded geostrophic EKE ($0.032$ m$^2$ s$^{-2}$).

\section{Bulk statistics of the surface lateral velocity gradient tensor}
To study the seasonality in the surface velocity, we calculate the lateral velocity gradient tensor
\begin{equation}
\left[\begin{matrix} u_x & u_y\\ v_x&v_y \end{matrix}\right]
\end{equation}
using a centered
second-order finite difference scheme --- we obtain qualitatively similar results
using spectral differentiation of velocity fields periodized with reflections;
statistics of fields obtained via spectral differentiation are 10-15$\%$ larger.
We then diagnose
the vertical vorticity
\begin{equation}
\zeta \equiv v_x - u_y\, ,
\end{equation}
lateral rate of strain
\begin{equation}
  \alpha \equiv [(u_x-v_y)^2 + (u_y + v_x)^2]^{1/2}\,,\\
\end{equation}
and horizontal divergence
\begin{equation}
\delta \equiv u_x + v_y\, .
\end{equation}
Because vortices, fronts, and IGWs are ubiquitous at submesoscales, second-order
statistics of vorticity, strain, and divergence highlight these
fine lateral scales \citep[e.g.,][]{shcherbina_etal2013}.

The model solutions depict seasonality in vorticity (Figures \ref{fig1}b-c): large values of
fine-grained vertical vorticity are observed in early spring with maximum
values as large as $4f$, where $f$ is
the local planetary vorticity, and root-mean-square (RMS) of about $0.4f$. In early
fall, the situation is the opposite: the vertical vorticity is relatively coarse-grained;
its local maximum and RMS are both smaller than $0.5f$.
Indeed, the vorticity and rate of strain are strongest in wintertime (Figure \ref{fig2}):
in both simulations, the RMS vorticity and strain rate are about twice as large in
late winter/early spring
than in late summer/early fall. Because the wintertime vorticity and strain rate
are dominated by the smallest scales in the flow (the KE spectrum is shallower than
a $-3$ power law in winter), increasing the resolution from
1/24$^\circ$ to 1/48$^\circ$ increases the wintertime RMS vorticity
and strain by about 40$\%$.

@@

Subinertial flows account for the bulk of vertical vorticity and strain rate
($T_{32.5} = 2\pi/f_{32.5}\approx 22.3$h, where $f_{32.5}$ is the inertial frequency at
the  mid-latitude $32.5^\circ$N): daily-averaging the velocity fields suppresses super-inertial
motions and reduces the RMS
vorticity by 40$\%$ and the RMS strain by 10$\%$; the seasonal
cycle remains strong (see
dashed lines in figures \ref{fig2}a-b). Indeed, submesoscale flows account for
most of this seasonal cycle: smoothing the velocity fields
with a Hanning filter with cut off scale of 100 km dramatically reduces the RMS
vorticity and strain rate. The reduction in variance is about 80$\%$ in winter,
yielding RMS vorticity and strain  rate roughly consistent with the diagnostics from
AVISO gridded geostrophic velocities (compare red lines to black lines in figures \ref{fig2}a-b).
The picture that emerges is consistent with recent studies: shallow baroclinic
instabilities energize the submesoscales in winter, drawing from the available
potential energy stored in large lateral buoyancy gradients in deep mixed layers \citep{sasaki_etal2014,callies_etal2015,callies_etal2016}.

The seasonal cycle of the horizontal divergence, however, showcases the complexity
of the upper-ocean annual variability. If submesoscale turbulence dominated
the near-surface variability all year, then the seasonal cycle of horizontal divergence,
vertical vorticity, and lateral strain rate would be in phase \citep[e.g.,][]{sasaki_etal2014}.
While there is a clear wintertime peak in divergence of daily-averaged velocity
(see dashed lines in figure \ref{fig2}c; RMS divergence $\sim0.1 f$ in the 1/48$^\circ$
simulation), the hourly fields show
a stronger enhancement of
lateral divergence in late summer/early fall (RMS divergence $\sim0.22 f$ in the 1/48$^\circ$
simulation). Because the 1/48$^\circ$ simulation better resolves
smaller-scale submesoscale flows, a secondary RMS divergence peak in winter  is
nearly as strong as in summer. Submesoscale fronts and eddies evolve
relatively fast, and there is no clear
temporal and spatial scale separation between those motions and IGWs
\citep[][]{mcwilliams2016}:
daily-averaging the velocity fields efficiently suppresses the summertime horizontally
divergent flows, but also reduces the wintertime lateral divergence by about 50$\%$.
Figure \ref{fig2}c also shows that most of the lateral divergence is associated
with submesoscale flows: smoothing the velocity fields with a 100-km-cutoff
suppresses more than 80$\%$ of the RMS divergence.

\begin{figure*}[ht]
  \begin{center}
    \includegraphics[width=.8\textwidth]{figs/fig3.pdf}
 \caption{\small Seasonal variation of joint probability distributions of
          surface lateral velocity gradient tensor:  vorticity vs.\
          strain rate (a through d),
          and vorticity vs.\ Laplacian of sea-surface height (e through h) in April (a, b,
          e, f) and October (c, d, g, h).
          Dashed lines in (a) through (d) represent one dimensional shear flow $\alpha = \pm\zeta$,
          characteristic of fronts. Dashed lines in (e) through
          (h) represent geostrophic flow $\zeta = \frac{g}{f}\nabla_h^2 \eta$.}
 \label{fig3}
 \end{center}
\end{figure*}

@@

The April vorticity-strain joint-PDF has a shape characteristic of submesoscale turbulence
 (Figures \ref{fig3}a-b). The alignment of vorticity and strain $\alpha \sim \pm\zeta$
 with strong positive skewness are fingerprints of submesoscale fronts
 \citep{shcherbina_etal2013,mcwilliams2016}. The shape of the vorticity-strain
 joint-PDF is similar for hourly and daily-averaged fields, although
 the vorticity skewness reduces from 1.4 to 1.13 from hourly to daily-averaged.
 The April results indicate that wintertime submesoscale
 surface velocity is strongly dominated by submesoscale turbulence.
 The daily-averaged fields
 are largely in  geostrophic balance,
but the hourly velocity and sea-surface height fields have an important
ageostrophic component as depicted by the joint-PDF of vorticity-Laplacian of SSH
winter (Figure \ref{fig3}d-e).

@@

\begin{figure*}[ht]
  \begin{center}
    \includegraphics[width=.9\textwidth]{figs/fig4.pdf}
 \caption{\small LLC4320 wavenumber-frequency spectrum of SSH variance in (a) April and (b) October.
          (c) Wavenumber spectrum of SSH variance --- the integral of (a) and (b) over frequency.
          In (a) and (b), the light gray shaded region depicts the dispersion relations for inertia-gravity
          waves from mode 1 through mode 4 across the latitudinal domain; horizontal dashed lines represent the semi-diurnal
          lunar tidal frequency ($M_2$), and the inertial frequency at mid-latitude
          ($f_{32.5}$) and its first harmonic (2$f_{32.5}$). }
  \label{fig4}
 \end{center}
\end{figure*}

Mesoscales flows and the lunar semi-diurnal tide contain most of the
SSH variance (Figures \ref{fig4}a-b). But there is significant SSH variance at
submesoscales, both sub-inertial
and super-inertial. Sub-inertial SSH variance at submesoscales is larger in April, whereas
super-inertial SSH variance at submesoscales is larger in October. At scales smaller
than 100 km (submesoscales), the SSH variance is 66$\%$ larger in October than in April.
IGWs, particularly modes two through four, account for most of the SSH variance
increase in summer --- higher vertical
modes are more sensitive to changes in near-surface stratification (supporting information).
In October, super-inertial IGWs account for a SSH standard deviation of about 4.9 cm,
comparable to the April sub-inertial SSH standard deviation of $\sim$5.2 cm (Table \ref{tab1}).
The
Similarly, the submesoscale super-inertial surface KE is 63$\%$ larger in October
than in April (supporting information).

@@

Figure \ref{fig4}c shows the integral over frequency
of wavenumber-frequency spectra in figures \ref{fig4}a-b.
Both in April and October, the
SSH wavenumber spectrum approximately follows a -2 power-law at submesoscales
(solid lines in figure \ref{fig4}c)  --- there is
small difference between the spectra owing to the phase cancellation between the
seasonal cycle of submesoscale turbulence and inertia-gravity waves. The sub-inertial flow
has steeper submesoscale SSH variance wavenumber spectrum (dashed lines in figure \ref{fig4}c).
 The difference between the SSH variance wavenumber spectrum for all frequencies and
 sub-inertial frequencies is dramatic in October when super-inertial flows account for
 about 80\% of the SSH variance and KE at submesoscales --- we obtain similar results by computing
SSH variance and KE wavenumber spectra directly from hourly and daily-averaged
SSH velocity fields.

@@

Internal tide energy flux does not show a seasonal modulation \citep[e.g.,][]{alford2003};
hence one expects a strong seasonality in the near-surface
expression of internal tides and
higher frequency IGWs if the near-surface stratification varies strongly.
The intermittent windy near-inertial wave (NIW) generation peaks in early winter, but
NIW energy has a similar energy level in October and April
\citep{alford_etal2016}. Furthermore, in our simulations the higher frequency inertia-gravity-wave
account for most of the seasonality in super-inertial surface KE and SSH
(Figure \ref{fig4}a-b).

The seasonally varying mixed-layer depth challenges observation and simulation of
flows at mixed-layer deformation-radius length scale in observational programs and models with
fixed resolution \citep{buckingham_etal2016}. But the mixed layer available potential energy
is lowest in summer owing to shallow mixed-layer depths and weak lateral buoyancy
gradients \citep{callies_etal2015}. Moreover, in baroclinic turbulence, the deformation
radius has only a secondary, catalytic role --- most of the baroclinic convertion occurs at
larger scales \citep{larichev_held1995,callies_etal2016}.
Thus, the model failure to resolve the summertime
ML deformation radius is unlikely to account for the vigorous seasonality in submesoscale
turbulence.

In the context of SWOT, these results imply that if one is interested
only in the geostrophic flow, so that surface velocity can be diagnosed
from SSH, then the noise-to-signal ratio will have a strong seasonality owing
to the presence of IGWs --- the coherent fraction of internal tides may be removed
efficiently.
An important caveat is that the Kuroshio Extension may be typical of a mesoscale-rich
subtropical regions, but it is unlikely to be
representative of other regions such as low-eddy-kinetic-energy eastern boundary
currents.  We plan to report on the geographic variability of submesoscale
seasonality in a future study.
hing.
